---
title: Importing data
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
  github_document: default
---

```{r global_options, include=TRUE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE, cache=FALSE,
                      warning=FALSE, message=FALSE,
                      cache.lazy = FALSE)

## ## Setup the cache and file directory
## origCache <- knitr::opts_chunk$get("cache.path")
## base <- sub("_cache/.*$", "", origCache)
## knitr::opts_chunk$set(fig.path = file.path(getwd(), "data", base, 'figures/'))

## Load packages
library(tidyverse)
library(knitr)
library(here)

## Setup some plotting details
coul <- colorRampPalette(RColorBrewer::brewer.pal(8, "RdYlBu"))(25)
source("helpers.R")
```
 
```{r directories}
## ## Common project name
## outputdir = file.path("data", base)
## if(!dir.exists(outputdir)) dir.create(outputdir)


base = "00-covariates"
here::i_am("00-covariates.Rmd")
knitr::opts_chunk$set(fig.path = here::here("data", base, 'figures/'))
outputdir = here::here("data", base)
if(!dir.exists(outputdir)) dir.create(outputdir)
```


There are three datasets for this project.

* Climatology data, from !()[], placed here:

* Non-climatology data, from ~()[], placed here:

* Depth data, downloaded from CMAP, placed here:


# Climatology data

The **climatology** chlorophyll dataset is a monthly dataset having averaged 20
years worth of data for each month (Jan-Dec).

There are multiple resolutions of this data; we mainly make use of the 2-degree resolution data.

```{r climatology}
## Read in data
datadir = "/home/sangwonh/Dropbox/research/usc/ocean-provinces/omd/data"
filenames = c("tabulated_darwin_montly_clim_045_090_ver_0_2_6.csv",
              "tabulated_darwin_montly_clim_090_180_ver_0_2_6.csv",
              "tabulated_darwin_montly_clim_180_360_ver_0_2_6.csv",
              "tabulated_darwin_montly_clim_360_720_ver_0_2_6.csv",

              "tabulated_geospatial_montly_clim_045_090_ver_0_2_5.csv",
              "tabulated_geospatial_montly_clim_090_180_ver_0_2_5.csv",
              "tabulated_geospatial_montly_clim_180_360_ver_0_2.csv", ## The 2.5 version is missing.
              "tabulated_geospatial_montly_clim_360_720_ver_0_2_5.csv")
ddat = read.csv(file.path(datadir, filenames[2])) %>% as_tibble()
rdat = read.csv(file.path(datadir, filenames[2+4])) %>% as_tibble()
```

# Non-climatology data

Non-climatology dataset was compiled by Bror Johnsson and shared as netcdf
files.

```{r nonclimatology}
print('Hi world')
```

# Depth data

The depth dataset has been downloaded from Simon's CMAP, from the tables "" and
"". Here is the code to do compile this data.

```{r}
```
