---
title: "Depth analysis: Earthmovers distance on Chlorphyll"
author: "Sangwon Hyun, Jacob Bien"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=14,
                      fig.height=5, echo=TRUE, warning=FALSE, message=FALSE, eval=TRUE, cache=FALSE)
knitr::opts_chunk$set(cache.path = "rmd-cache/hawaii-3-18-2020/")

## Read in libraries
## library(Matrix)
## library(transport)
## library(rworldmap)
## library(viridis)
## library(maps)
library(omd)
## library(lattice)
library(cmap4r)
library(dplyr)
```

First, colocalize the Darwin data with the HOT data.

```{r colocalize}
## Set a location for the plots
outputdir = file.path("~", "depthplots")
dir.create(outputdir, recursive = TRUE)

## Colocalize the data
initialize_cmap(cmap_key = "5e05c500-d68d-11e9-9d3b-4f83fcec4710")
df = get_metadata('tblHOT_Bottle', 'HPLC_chl3_bottle_hot')
dat = compile(sourceTable='tblHOT_Bottle',
              sourceVar='HPLC_chla_bottle_hot',
              targetTables="tblDarwin_Ecosystem",
              targetVars="CHL",
              dt1="2013-03-03",
              dt2="2015-04-03",
              ## lat1=21.847,
              ## lat2=22.75,
              ## lon1=-158.363,
              ## lon2=-157.9,
              lat1 = df$Lat_Min,
              lat2 = df$Lat_Max,
              lon1 = df$Lon_Min,
              lon2 = df$Lon_Max,
              ## depth1=0.3,
              ## depth2=4813.8,
              depth1 = df$Depth_Min,
              depth2 = df$Depth_Max,
              temporalTolerance=c(3),
              latTolerance=c(1),
              lonTolerance=c(1),
              depthTolerance=c(20))
saveRDS(dat, file= file.path(outputdir, "dat-higher-depth-tol.RDS"))
dat = readRDS(file= file.path(outputdir, "dat-higher-depth-tol.RDS"))

## See a summary of this data.
dat %>% summary() %>% print()
dat %>% dim() %>% print()
```

Now, make plots of CHL from the two data sources, over time:

```{r make-plots, animation.hook='ffmpeg', dev='jpeg', interval=2, ffmpeg.format="gif", fig.width=15, fig.height=7}
## Make plots for each
for(ii in 1:88){
  onetime = dat$time %>% unique() %>% .[ii]
  onedat = dat  %>% select(lat, lon, time, CHL, depth, HPLC_chla_bottle_hot)
  
  ## Also isolate to a single latitude
  onedat = onedat %>% filter(time == onetime)
  onedat = onedat %>% arrange(depth)

  if(nrow(onedat) == 0) next

  ## Make two plots
  png(file = file.path(outputdir, "overlaid", paste0("depthplot-", ii, ".png")),
                       width = 800, height = 500)
  par(mfrow=c(1,2), cex=1.2, oma = c(1,1,4,1))

  ## First plot is Darwin data
  if(any(!is.na(onedat$CHL))){
    y = onedat$CHL
    y = y/max(y, na.rm=TRUE)
    x = onedat$depth
    plot(y = y, x = x,
         type='o', cex = 1.5, log="x",  xlab = "Depth", ylab = "Darwin CHL", col='grey50', lwd=3, pch=16)
    mtext(onetime, side = 3, outer=TRUE, cex=2)
  } else {
    plot(NA, ylim=c(0,1))
  }
  title(main = "Darwin Chl-A (black) and real data  vs   depth")

  ## Add real data to first plot
  if(any(!is.na(onedat$HPLC_chla_bottle_hot))){

    ## Add an overlaid line
    y = onedat$HPLC_chla_bottle_hot
    y = y / max(y, na.rm=TRUE)
    x = onedat$HPLC_chla_bottle_hot
    lines(y=y, x=x, col = 'red', lwd=3, pch=16, type='o')

    ## Also add scatterplot
    plot(y=onedat$CHL, x=onedat$HPLC_chla_bottle_hot, pch=16, ylab = "Darwin CHL", xlab = "HOT CHLA", col='grey50', cex=1.5,
          type='p')
    title(main="Scatterplot between two sources")
  } else {
    plot(NA, ylim=c(0,1))
    title(main="Scatterplot between two sources")
  }
  graphics.off()
}
```

# Other data leads

The Darwin data is here:
https://simonscmap.com/catalog/datasets/Darwin_Ecosystem

The HOT data is here:
https://simonscmap.com/catalog/datasets/HOT_Bottle_ALOHA

Also a different data that also contains CHL-A data, but from a different source.
https://simonscmap.com/catalog/datasets/HOT_PP

Chris Follet strongly recommends CTD data:

"My first instinct is to try and use depth profiles from Station ALOHA, focusing
on the fluorescence derived Chlorophyll from the CTD drops. There is also some
very high resolution ‘Wire Walker’ Data for a handful of cruises. I think that
both of these data types are in CMAP. My recollection is that Aditya built an R
package to deal with CMAP so this is probably easy to get/use? If this is so,
there is a very clean way to start using the DARWIN model CHL which is also in
CMAP https://simonscmap.com/catalog/datasets/Darwin_Ecosystem at 3 day depth
resolution. (I added Stephanie to the chain in case she has thoughts here)."

"I would recommend also using the CTD data (realtime sensor dropped down the
wire: chl from fluorescence, but there is also Salinity, Temperature, densitiy,
and nitrate which can be compared directly to DARWIN) because it is of much
higher depth resolution. I think any precision lost using fluorescence is more
than gained back in the shape resolution. I am very sure this data is also in
CMAP, but to make sure you actually receive an email reply from me the data
website from HOT is
https://hahana.soest.hawaii.edu/hot/hot-dogs/cextraction.html which should help
in finding the correct table in CMAP."

 
I *think* this is the right CTD dataset:
https://simonscmap.com/catalog/datasets/HOT_CTD


# Early impressions

A preliminary look at the data seems to suggest that OMD is perfect in
explaining the difference. A scatterplot of Chlorophyll by depth seems to show
some *mass shifts* along depth axis that isn't explained well by:

1. Scatterplots between the CHL values, or
2. An L2 distance.

The next steps are to (1) colocalize + download CTD data, and (2) visual
comparisons (3) then, apply OMD.
