---
title: "Data experiments (Continue from Jan 20th 2021)"
author: "Sangwon Hyun, Jacob Bien, and the Ocean Provinces working group"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height=7, echo=TRUE, warning=FALSE,
                      message=FALSE, eval=TRUE, cache=FALSE)
knitr::opts_chunk$set(cache.path = "rmd-cache/feb2-2021-experiments/")
```

# Read in libraries

```{r}
## Read in libraries
library(latticeExtra)  
library(Matrix) 
library(transport)
library(dplyr)
library(omd)
library(raster) 
## destin = "~/Dropbox/research/usc/ocean-provinces/omd/" ##
## la("omd") ## Import from destin
```


# Summary

Last time, we saw how experiments might apply to a checkerboard pattern
(global + local change).

Also, in `feb2-2021.Rmd`, we saw improvements in calculating Sinkhorn.

In this document, we conduct some experiments on the checkerboard pattern example.

(One residual question was: Why does transport go to 1 (and not 0) as the image
is coarsened? <-- solved)

# Experiments 

This is what smoothing does to the two mean matrices:

```{r, fig.width=10, fig.height=5}
n = 2^5 
M_mean = matrix(0, ncol = n, nrow=n)
sig = 1 
M1_mean = M_mean %>% add_global(2) %>% add_checkerboard(4, sig)  %>% add_checkerboard(8, sig) 
M2_mean = M_mean %>% add_global(2)
M1_mean = M1_mean + 5 
M2_mean = M2_mean + 5

## No smoothing, to highly smoothed means
for(d in c(2,5,10, 15, 20, 22, 24, 26, 30, 30, 31)){
  gridExtra::grid.arrange(M1_mean %>% smoothmat(d) %>% drawmat_precise(main = "Image 1"),
                          M2_mean %>% smoothmat(d) %>% drawmat_precise(main = "Image 2"),
                          ncol = 2,
                          top = paste0("Smoothing window size = ", d, "x", d))
} 
```

Now, visualizing two things, over a signal size (`sig`, which governs how prominent the checkerboard pattern is):

1. The difference in the two mean matrices (no noise), generated as:
     + Take mean matrix 1 and 2.
     + Coarsen both by a window size of (d x d).
     + Measure the average of absolute pixel-wise difference.

2. The OMD between the two noisy images:
     + Take mean matrix 1 and 2.
	 + Add N(0, 0.1) random noise to all pixels.
     + Coarsen both noisy images by a window size of (d x d).
     + Measure the average of absolute pixel-wise difference.



```{r, fig.width=10, fig.height=5}
for(sig in c(0, 0.1, 0.5, 1, 2)){

  ## Generate mean  
  M1_mean = M_mean %>% add_global(2) %>% add_checkerboard(4, sig)  %>% add_checkerboard(8, sig) 
  M2_mean = M_mean %>% add_global(2)
  M1_mean = M1_mean + 5
  M2_mean = M2_mean + 5
  
  ## Setup 2-column plots
  par(mfrow = c(1, 2), oma = c(1,1,3,1))
  sizes = seq(from = 1, to = 32, by=1)
  diff_in_means = sapply(sizes, function(d){
    mean(abs(range(smoothmat(M1_mean, d) - smoothmat(M2_mean, d))))
  })
  diff_in_means %>% plot(type = 'o', x = sizes, xlab = "Window size", ylab = "avg. of absolute pixel-wise differences")
  abline(v = c(4, 8), lty = 2)
  
  ## Generate noisy data
  sigma = .1
  nsim = 16
  
  ## Now, as a function of coarsening, we want to see the trend of OMD.
  dists_list = list() 
  sizes = seq(from=1, to=32, by=1)
  for(ii in 1:length(sizes)){
    d = sizes[ii]
    dists = mclapply(1:nsim, function(isim){
      res <- omd((M1_mean + rnorm(length(M1_mean), 0, sigma)) %>% smoothmat(d),
                 (M2_mean + rnorm(length(M2_mean), 0, sigma)) %>% smoothmat(d),
                 type = "transport")
      return(res$dist)
    }, mc.cores = 4)
    dists_list[[ii]] = dists %>% unlist()
  }
  dists_list = lapply(dists_list, unlist)
  distmat = dists_list %>% do.call(rbind, .) %>% as.matrix()
  distmat %>% matplot(lwd=3, col=rgb(1,0,0,0.3),
                      xlab="Data smoothing window\n left=fine, right=coarse",
                      ylab = "OMD", pch=16,
                      x = sizes)
  
  apply(distmat, 1, mean) %>% lines(type='o', lwd=3, col='black',
                                      xlab="Data smoothing window\n left=fine, right=coarse",
                                      ylab = "OMD",
                                      x = sizes)
  abline(v = c(4, 8))
  legend("topright", pch = c(16, NA), lwd = c(NA, 3), lty = c(NA, 1), col = c(rgb(1,0,0,0.3), rgb(0,0,0)),
         legend = c("Each simulation", "Mean"))
  mtext(text = paste0("Size of global pattern = ", sig), outer = TRUE, side = 3, cex = 1.5)
}
```

# Next steps

- Use optimality conditions of the Sinkhorn convex optimization problem as a
  stopping rule.

- Investigate what's going on with the later (coarsen window 24-32) parts of the
  plots.

- Change the global pattern to also shift a bit.

- Make the explanation slightly clearer.

- Investigate the L1 Wasserstein Wavelet paper
  (https://ttic.uchicago.edu/~ssameer/Research/Papers/WEMD_CVPR08.pdf) and
  another Fourier-based method (https://arxiv.org/pdf/2005.06530.pdf). Christian
  will look into (Matlab) code that exists.

- (Send Bror the scripts, after cleaning a little bit)

- Visualize the transports.


<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

