---
title: "Learning more about Sinkhorn (Continue from Jan 20th 2021)"
author: "Sangwon Hyun, Jacob Bien, and the Ocean Provinces working group"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height=7, echo=TRUE, warning=FALSE,
                      message=FALSE, eval=TRUE, cache=TRUE)
knitr::opts_chunk$set(cache.path = "rmd-cache/feb2-2021/")
```

# Load libraries

```{r}
## Read in libraries
library(latticeExtra) 
library(Matrix)
library(transport)
library(dplyr)
library(omd)
## la("omd") ## Import from destin
## destin = "~/Dropbox/research/usc/ocean-provinces/omd/" ##
```

# Summary

Last time, we saw how experiments might apply to a checkerboard pattern
(global + local change).

In this document, we fix an error of mine (of discouraging mass from staying
put), and then improve Sinkhorn by changing the convergence criterion of the
algorithm to be relative.



# Data

We'll be comparing two images, from these mean images:

```{r, fig.width=5, fig.height=5}
## Generate means
n = 2^3
M_mean = matrix(0, ncol = n, nrow=n)
sig = 1
M1_mean = M_mean %>% add_global(2) %>% add_checkerboard(2, sig) 
M2_mean = M_mean %>% add_global(2)
M1_mean = M1_mean + 5
M2_mean = M2_mean + 5
stopifnot(all(dim(M1_mean)==dim(M2_mean)))
M1_mean %>% drawmat_precise()
M2_mean %>% drawmat_precise()
```

# Sinkhorn, visualized by iteration

We're stopping when maximum of row sums doesn't change that much (1E-14).

```{r, fig.width=7, fig.height=7}
set.seed(10)
sigma = .5
M1 = M1_mean + rnorm(length(M1_mean), 0, sigma)
M2 = M2_mean + rnorm(length(M2_mean), 0, sigma) 
res <- omd(M1, M2, type = "sinkhorn")

## Plot two quantities over iterations
par(mfrow = c(2, 1), mar = c(5,5,2,2))
gridcol = rgb(0,0,0,0.2)
na.omit(res$sinkhorn_object$gaps) %>% log10() %>% plot(type='l', pch=16, cex=.5, log="x",
                                                       xlab = "Sinkhorn iteration", ylab = "log10(row sum difference from \n previous iteration)")
abline(h=seq(from=-20, to=0, by = 1), col = gridcol)
abline(v=seq(from=0, to=100000, by = 100), col = gridcol)
abline(h=-14, lwd=2, col='red')

na.omit(res$sinkhorn_object$dists) %>%  plot(type='l', pch=16, cex=.5, log="x",
                                                       xlab = "Sinkhorn iteration", ylab = "OMD")
abline(h=seq(from=1E-15, to=1, by = 1E-2), col = gridcol)
abline(v=seq(from=0, to=100000, by = 100), col = gridcol)
```

# Sinkhorn with relative stopping

Instead of an absolute for the criterion (1E-15), how about if we used
stabilization of the distances? Cuts down on run time.

(Red line is absolute threshold for row sum difference across iterations)

(Blue line is time used by transport R package)

```{r}
sinkhorn_rel_eps = NULL

set.seed(10)
sigma = .5
M1 = M1_mean + rnorm(length(M1_mean), 0, sigma)
M2 = M2_mean + rnorm(length(M2_mean), 0, sigma) 

before_time = Sys.time()
transportR_dist <- omd(M1, M2, type = "transport")$dist
after_time = Sys.time()
transport_lapse_time = (after_time - before_time) %>% as.numeric() %>% round(3)

## Initialize container
sinkhorn_rel_eps_list =  c(NA, 1E-10, 1E-8, 1E-7, 1E-6, 1E-4)
reslist = vector(mode="list", length=6)
lapse_times = rep(NA, 6)

## Calculate Sinkhorn Distance
for(ii in 1:6){
  sinkhorn_rel_eps = sinkhorn_rel_eps_list[ii]
  before_time = Sys.time()
  if(is.na(sinkhorn_rel_eps)) sinkhorn_rel_eps = NULL
  res <- omd(M1, M2, type = "sinkhorn", sinkhorn_rel_eps = sinkhorn_rel_eps)
  reslist[[ii]] = res
  lapse_times[[ii]] = (Sys.time() - before_time) %>% as.numeric() %>% round(3)
}
```

```{r, fig.width=7, fig.height=14}
## Plot two quantities over iterations
par(mfrow = c(6, 2), mar = c(5,5,2,2))
sinkhorn_rel_eps_list =  c(NA, 1E-10, 1E-8, 1E-7, 1E-6, 1E-4)
for(ii in 1:6){
  gridcol = rgb(0,0,0,0.2)
  sinkhorn_rel_eps = sinkhorn_rel_eps_list[ii]
  res = reslist[[ii]]
  na.omit(res$sinkhorn_object$gaps) %>% log10() %>% plot(type='l', pch=16, cex=.5, log="x",
                                                         xlab = "Sinkhorn iteration", ylab = "log10(row sum difference from \n previous iteration)")
  abline(h=seq(from=-20, to=0, by = 1), col = gridcol)
  abline(v=seq(from=0, to=100000, by = 100), col = gridcol)
  abline(h=-14, lwd=2, col='red')
  
  na.omit(res$sinkhorn_object$dists) %>%  plot(type='l', pch=16, cex=.5, log="x",
                                               xlab = "Sinkhorn iteration", ylab = "OMD",
                                               ylim = c(0, transportR_dist))
  abline(h = transportR_dist, col = "blue", lwd=2)
  abline(h=seq(from=1E-15, to=1, by = 1E-2), col = gridcol)
  abline(v=seq(from=0, to=100000, by = 100), col = gridcol)
  legend("topleft", legend = paste("rel. stop = ", sinkhorn_rel_eps), cex=1.5, bty="n")

  ## measure time
  legend("bottomright", legend = paste0("run time = ", lapse_times[ii], "\n",
                                        "transportR run time = ", transport_lapse_time), cex=1.5, bty="n")
}
```

# Sinkhorn as a function of the regularization parameter

Next, as a function of lambda, how does this hold up? Let's see

- The (dis) agreement between Sinkhorn and software from `transport` R package.

- Run time.

```{r, fig.width=5, fig.height=10}
sigma = .5
nsim = 30
par(mfrow=c(4,1))
for(sinkhorn_lambda in c(5, 10, 20, 30)){
  dists = mclapply(1:nsim, function(isim){
    printprogress(isim, nsim)
    set.seed(isim * sinkhorn_lambda)
  
    ## Add noise
    M1 = M1_mean + rnorm(length(M1_mean), 0, sigma)
    M2 = M2_mean + rnorm(length(M2_mean), 0, sigma) 
    
    ## Calculate the two distances
    costm <- omd(M1, M2, type="transport")$costm

    before_time = Sys.time()
    transportR_dist <- omd(M1, M2, costm = costm, type="transport")$dist
    transport_time = (Sys.time() - before_time) %>% as.numeric() %>% round(3)

    before_time = Sys.time()
    sinkhorn_dist <- omd(M1, M2, costm = costm, type = "sinkhorn", sinkhorn_rel_eps = 1E-6, sinkhorn_lambda = sinkhorn_lambda)$dist
    sinkhorn_time = (Sys.time() - before_time) %>% as.numeric() %>% round(3)
  
    return(c(sinkhorn = sinkhorn_dist, transport = transportR_dist, sinkhorn_time = sinkhorn_time, transport_time = transport_time))
  }, mc.cores = 8)
  
  distmat = do.call(rbind, dists) %>% as_tibble()
  plot(y = distmat$transport,
       x = distmat$sinkhorn,
       xlab = "sinkhorn", ylab = "transport (exact)", log="xy")
  title(main = paste0("lambda = ", sinkhorn_lambda))
  distmat$sinkhorn_time %>% mean() %>% round(3) %>% paste0("sinkhorn time=", .) %>% legend("topleft", legend=., bty="n")
  distmat$transport_time %>% mean() %>% round(3)%>% paste0("transport time=", .) %>% legend("bottomright", legend=., bty="n")
  abline(0,1)

}
```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
