---
title: "Earthmovers distance and optimal transport on Chlorophyll (take 2)"
huthor: "Sangwon Hyun, Jacob Bien"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=14,
                      fig.height=5, echo=TRUE, warning=FALSE, message=FALSE, eval=TRUE, cache=TRUE)
knitr::opts_chunk$set(cache.path = "rmd-cache/hawaii-3-18-2020/")

## Read in libraries
library(Matrix) 
library(transport)
library(rworldmap)
library(viridis)
library(maps)
library(omd)
library(lattice)
```


```{r read-data}
## Read in data
datadir = "/home/sangwonh/Dropbox/research/usc/ocean-provinces/omd/data"
filenames = c("tabulated_darwin_montly_clim_045_090_ver_0_2.csv",
              "tabulated_darwin_montly_clim_090_180_ver_0_2.csv",
              "tabulated_darwin_montly_clim_180_360_ver_0_2.csv",
              "tabulated_darwin_montly_clim_360_720_ver_0_2.csv",
              "tabulated_geospatial_montly_clim_045_090_ver_0_2_5.csv",
              "tabulated_geospatial_montly_clim_090_180_ver_0_2_5.csv",
              "tabulated_geospatial_montly_clim_180_360_ver_0_2_5.csv",
              "tabulated_geospatial_montly_clim_360_720_ver_0_2_5.csv",
              "tabulated_geospatial_montly_clim_045_090_ver_0_2.csv",
              "tabulated_geospatial_montly_clim_090_180_ver_0_2.csv",
              "tabulated_geospatial_montly_clim_180_360_ver_0_2.csv",
              "tabulated_geospatial_montly_clim_360_720_ver_0_2.csv")
```

# Data region

We'll focus on the yellow box in the map.

```{r new-box}
dat = read.csv(file.path(datadir, filenames[1]))
mydat = dat[which(dat[,"month"]==1),]
colfun = colorRampPalette(c("blue", "red"))

pdf("../figures/climatology-map.pdf", width = 7, height=6.7)
par(mar=c(0,0,0,0))
map("world", fill=TRUE, col="white", bg="white", ylim=c(0, 70), mar=c(0,0,0,0),
    xlim = c(-180, -90))
## plot(newmap)##, xlim = c(-20, 59), ylim = c(35, 71), asp = 1)
centers = sample(mydat$Chl, 5)
points(mydat[ ,c("lon", "lat")], pch=15, cex=20,
       col=colfun(20)[as.numeric(cut(mydat$Chl,breaks = 20))])
points(mydat[ ,c("lon", "lat")], pch=15, cex=5,
       col=colfun(20)[as.numeric(cut(mydat$Chl,breaks = 20))])
points(mydat[ ,c("lon", "lat")], pch=15, cex=2.5,
       col=colfun(20)[as.numeric(cut(mydat$Chl,breaks = 20))])
map("world", fill=TRUE, col="white", bg="white", ylim=c(-90, 90), mar=c(0,0,0,0), add=TRUE)
## New box 
lonrange = c(-180, -135)
latrange = c(20, 45)
rect(lonrange[1], latrange[1], lonrange[2], latrange[2],
     border = c("black"), lwd=5)
legend("bottomleft", bty = "n", ## fill="blue", bg = "white", 
       legend = "Chlorophyll-A in January\n\n", cex=2)
graphics.off()
```

# Darwin (between adjacent months)

```{r all-pairs-of-months, eval=TRUE}
all.mats = list()
all.pmats = list()
for(ii in 1:12){
  ## One month's data
  mydat = dat[which(dat[,"month"]==ii),]
  one.dat = mydat[which(lonrange[1] < mydat[,"lon"] & mydat[, "lon"] < lonrange[2] &
                        latrange[1] < mydat[,"lat"] & mydat[, "lat"] < latrange[2]),]
  one.mat = make_mat(one.dat)
  one.mat = one.mat/sum(one.mat)
  all.mats[[ii]] = one.mat
  all.pmats[[ii]] = pgrid(one.mat)
}
```

Now, we obtain all twelve months' Darwin data and visualize optimal transports
in **adjacent** months (Jan to Feb, Feb to March, and so on). Note, I highlighed
the top 10% high-mass transfers in red.

```{r all-pairs-of-months-transport, fig.width=20, fig.height=60}
par(mfrow=c(6,2))
for(ii in 1:11){
  pgrid1 = pgrid(all.mats[[ii]])
  pgrid2 = pgrid(all.mats[[ii+1]])
  res <- transport(pgrid1, pgrid2, p=2,
                   method="aha")


  ## nbin = 2
  ## cutoffs = seq(from = min(res$mass), to = max(res$mass), length = nbin+1)
  cutoffs = quantile(res$mass, probs = c(0.9,1))
  jj = 1
  massrange = cutoffs[jj:(jj+1)]
  cols = rep(rgb(0,0,1,0.2), length(res$mass))
  cols[which(massrange[1] < res$mass & res$mass < massrange[2])] = rgb(1,0,0,0.7)

  plot(pgrid1, pgrid2, res,
       ## Style options
       mass="thickness", acol=cols,
       ## lwd=0.1,
       rot=TRUE, length=0.2)

  title(main = paste0( month.abb[ii], " to ", month.abb[ii+1]), cex.main=2)
}
```



# Earth Mover's Distance Matrix

Now, let's form a 24 x 24 distance matrix, which encodes *all* the pairwise
earth mover's distances between all months in real data and Darwin data.


```{r all-dist}
## DARWIN
all.pmats.darwin = list()
dat = read.csv(file.path(datadir, filenames[2]))
nmonth = 12
for(ii in 1:nmonth){
  ## One month's data
  mydat = dat[which(dat[,"month"]==ii),]
  one.dat = mydat[which(lonrange[1] <= mydat[,"lon"] & mydat[, "lon"] <= lonrange[2] &
                        latrange[1] <= mydat[,"lat"] & mydat[, "lat"] <= latrange[2]),]
  one.mat = make_mat(one.dat)
  ## one.mat = one.mat[-nrow(one.mat),]
  one.mat = one.mat/sum(one.mat, na.rm=TRUE)
  one.mat[which(is.na(one.mat))] = 0
  all.pmats.darwin[[ii]] = pgrid(one.mat)
}

## REAL
dat = read.csv(file.path(datadir, filenames[6])) ## real
all.pmats.real = list()
for(ii in 1:nmonth){
  ## if(ii==1) jj=2 ## replace this...................
  ## if(ii==12) jj=11
  ## if(ii %ni% c(1,12)) jj=ii

  ## One month's data
  mydat = dat[which(dat[,"month"]==ii),]
  one.dat = mydat[which(lonrange[1] <= mydat[,"lon"] & mydat[, "lon"] <= lonrange[2] &
                        latrange[1] <= mydat[,"lat"] & mydat[, "lat"] <= latrange[2]),]
  one.mat = make_mat(one.dat)
  ## if(ii!=12) one.mat = one.mat[-nrow(one.mat),]
  one.mat = one.mat/sum(one.mat, na.rm=TRUE)
  ## one.mat[which(is.na(one.mat))] = 0
  all.pmats.real[[ii]] = pgrid(one.mat)
}

all.pmats = c(all.pmats.real, all.pmats.darwin)
nmonth = 24
distmat = matrix(NA, nmonth, nmonth)
## mclapply(1:(nmonth)^2, function(imonth){
##   ii = imonth %% 12 + 1
##   jj = imonth - 12 * (ii-1)
for(ii in 1:nmonth){
  for(jj in 1:nmonth){
    if(ii>jj){
      res <- transport(all.pmats[[ii]], all.pmats[[jj]], p=2, method="aha")
      dist = wasserstein(all.pmats[[ii]], all.pmats[[jj]], tplan=res)
      distmat[ii,jj] = dist
      distmat[jj,ii] = dist
    }
  }
} 
## }, mc.cores=8)

diag(distmat) = 0 
```

```{r all-dists-quick-viz, fig.height=10, fig.width=10}
diag(distmat) = NA
colnames(distmat) = rownames(distmat) = c(paste0("r", 1:(nmonth/2)),
                                          paste0("d", 1:(nmonth/2)))
## image(Matrix(distmat))  
drawmat_precise(distmat, contour=FALSE, main="All months' EMD from Real & Darwin data",
                 panel = function(...){
                   panel.levelplot(...)
                   lattice::panel.abline(v = 12.5, lwd=3)
                   lattice::panel.abline(h = 12.5, lwd=3)
                 })
## abline(v=13, lwd=5)
## knitr::kable(as.data.frame(signif(distmat,3)))
```

We can take a look at certain distances; from left to right in the boxplot.

1. Between Darwin and Real.
2. Among Darwin data.
3. Among Real data.
4. Among Darwin data, adjacent months.
5. Among Real data, adjacent months.


**Compared to the larger white box before** we can see that the distances
between Darwin data and real data (left-most box) is even more separated from
the rest of the boxplots.

```{r all-dists-boxplot, fig.height=7, fig.width=15}
diag(distmat) = 0 
## Extract certain things
pairdists = sapply(1:12, function(ii){
  distmat[ii, ii+12]
})
distmat.real = distmat[1:12, 1:12]
distmat.darwin = distmat[13:24, 13:24]

## Combine them into a list
dists = list(pair= pairdists,
             darwin = distmat.darwin[lower.tri(distmat.darwin)],
             real = distmat.real[lower.tri(distmat.real)],
             darwin_serial = unlist(diag(distmat.darwin[-1,-ncol(distmat.darwin)])),
             real_serial = unlist(diag(distmat.real[-1,-ncol(distmat.real)])))

## Clean
dists = lapply(dists, function(a){
  if(any(is.na(a))){
    return(a[-which(is.na(a))])
  } else {
    return(a)
  }
})

names(dists)=c("darwin-to-real","darwin", "real", "darwin m2m", "real m2m")
boxplot(dists,
        col=c(rgb(0,0,1,0.2), rep(rgb(1,0,0,0.2),2), rep(rgb(0,1,0,0.2),2)),
        main="Distributions of Earth mover's distances", cex.main=2, cex.lab=2)
```


# Clustering and Multidimensional Scaling

From a distance matrix containing all pairwise distances, we can produce more
interesting analyses and visualizations.

**Hierarchical clustering** from the pairwise distances: at the beginning of the
process, each element is in a cluster of its own. The clusters are then
sequentially combined into larger clusters until all elements end up being in
the same cluster.



```{r all-dists-viz, fig.width=10, fig.height=6}
colnames(distmat) = rownames(distmat) = c(paste0("real-", 1:(nmonth/2)),
                                          paste0("darwin-", 1:(nmonth/2)))
distmat <- as.dist(distmat, diag = TRUE)

labelCol <- function(x) {
  if (is.leaf(x)) {
    ## fetch label
    label <- attr(x, "label") 
    ## set label color to red for A and B, to blue otherwise
    attr(x, "nodePar") <- list(lab.col=ifelse(sapply(label, substr, 1,4) == "real", "blue", "red"))
  }
  return(x)
}

## plot(hclust(distmat), axes=FALSE)
hc = hclust(distmat)
d <- dendrapply(as.dendrogram(hc), labelCol)
plot(d, axes=FALSE)
```

**Multidimensional scaling (MDS)** is a *dimension reduction* technique to
translate + visualize the level of similarity of $n$ individual cases of a
dataset. MDS translates pairwise 'distances' among a set of $n=24$ objects into
a configuration of $n=24$ points mapped into a lower-dimensional Cartesianspace
(e.g. 2D) while preserving the pairwise distances.

<!-- Basically, the coordinates of the 2d graph are in the eigenspace spanned by the -->
<!-- first two eigenvectors of the distance matrix. -->

```{r, fig.width=10, fig.height=10}
## Multidimensional scaling
fit <- cmdscale(distmat, eig=TRUE, k=2) # k is the number of dim

# plot solution
x <- fit$points[,1]
y <- fit$points[,2]
par(pty="s")
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",
     main="Metric MDS", type="n")
text(x, y, labels = c(paste0(1:(nmonth/2)),
                      paste0(1:(nmonth/2))),
     col =  c(rep("blue", nmonth/2), rep("red", nmonth/2)),
     cex=2)
legend("topright", fill=c('blue', 'red'), legend = c("real", "Darwin"))
```

Each data source is closer to each other, more so than the same month from the
two data sources.



# Some subtleties

What to do for optimal transport in the presence of land mass to obstruct
movement (water can't cross a peninsula)? One can simply modify the distances so
that, the Euclidean distance is the default, but whenever the arrow meets a land
mass, the distance is positive infinity so that it incurs an infinite cost for
any mass to cross a landmass. (If it really needs to, it will go around the
coast.)

Also, the Wasserstein's distance metric does not seem to be comparable between
different size objects. In other words, 
